---
title: "MoralFoundationsSurvey2"
output: 
  html_document:
    toc: true
date: "2023-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare the data

Load separate files for the positive and negative responses

```{r data loading, message = FALSE}
library(readr)

url_neg = "https://raw.githubusercontent.com/NicoleNisbett/Moral-Foundations/main/NEG_coded.csv"
neg_data <- read_csv(url(url_neg))

url_pos = "https://raw.githubusercontent.com/NicoleNisbett/Moral-Foundations/main/POS_coded.csv"
pos_data <- read_csv(url(url_pos))
```

Create the value scales based on TwIVI
```{r}

create_scales = function(file){
  data = file
  
  #creating openness to change scales
  subset1 <- data.frame(data$Q3_3,data$Q5_4,data$Q5_3,data$Q4_7)
  data$conformity_val <- rowMeans(subset1)
#Schwartz_5 Schwartz_6 Schwartz_15 Schwartz_16
  subset2 <- data.frame(data$Q5_5,data$Q5_2,data$Q3_4,data$Q4_3)
  data$self_direction_val <- rowMeans(subset2)
  data$SCH_Conservation <- data$conformity_val - data$self_direction_val
  
  #creating Self-enhancement/Self-trancendencies scales
  subset3 <- data.frame(data$Q3_2,data$Q3_7,data$Q4_4,data$Q5_6)
  data$universalism_val <- rowMeans(subset3)
  #
  subset4 <- data.frame(data$Q3_6,data$Q4_2,data$Q4_5,data$Q5_1)
  data$achievement_val <- rowMeans(subset4)
  data$SCH_SelfTrancendence <- data$universalism_val - data$achievement_val
  
  return(data)
}

negative = create_scales(neg_data)
negative$sentiment = "negative"

positive = create_scales(pos_data)
positive$sentiment = "positive"


```

## Descriptive Analysis

Rename some of the questions for ease and combine them
```{r}
names(positive)[names(positive) == "Q17"] <- "Income"
names(negative)[names(negative) == "Q17"] <- "Income"

names(positive)[names(positive) == "Q13"] <- "Age"
names(negative)[names(negative) == "Q13"] <- "Age"

names(positive)[names(positive) == "Q19_1"] <- "Political_Leaning"
names(negative)[names(negative) == "Q19_1"] <- "Political_Leaning"

combined = rbind(negative, positive)
```

Plot the distiribution for income
```{r}
library(ggplot2)
ggplot(combined, aes(x = Income))+ geom_bar() +scale_x_discrete(name = "Annual Household Income (in £1000s)", limits = c("<20", "20 - 40", "40 - 60", "60 - 80", "80 - 100", ">100"))
```

Distribution of educational background
```{r}
ggplot(combined, aes(x = Q15))+ geom_bar() + scale_x_discrete(name = "Highest Qualification", limits = c("No formal", "GCSE", "A-level", "Vocational", "Bachelors", "Masters", "Doctoral")) 

```


## Confirmatory Factor Analysis (CFA)
```{r message=FALSE, warning=FALSE}
library(lavaan)
set.seed(56789)
```

Get just the negative questions
```{r}
QsN= data.frame(negative$Q6_1_a, negative$Q6_2_a,negative$Q6_3_a,negative$Q6_4_a, negative$Q6_5_a,negative$Q6_6_a, negative$Q7_1_a, negative$Q7_2_a, negative$Q7_3_a, negative$Q7_4_a, negative$Q7_5_a, negative$Q7_6_a , negative$Q8_1_a, negative$Q8_2_a, negative$Q8_3_a, negative$Q8_4_a, negative$Q8_5_a, negative$Q8_6_a) 

names(QsN) = c("comp1", "comp2", "comp3", "fair1", "fair2", "fair3", "pur1", "pur2", "pur3", "auth1", "auth2", "auth3", "loy1", "loy2", "loy3", "anc1", "anc2", "anc3")

```

Create the model for the CFA based on all statements and run the model

```{r}
items = '
compassion =~ comp1 + comp2 + comp3
fairness =~ fair1 + fair2 + fair3
purity =~ pur1 + pur2 + pur3
authority =~ auth1 + auth2 + auth3
loyalty =~ loy1 + loy2 + loy3 
ancestors =~ anc1 + anc2 + anc3
'
#decent model but can be improved
cfa_modN = cfa(items, data = QsN)
summary(cfa_modN, standardized = TRUE, fit.measures=TRUE) 

```

Pur1 and Loy1 have low loading so remove them and re-run the model

```{r}
itemsN = '
compassion =~ comp1 + comp2 + comp3
fairness =~ fair2 + fair3
purity =~ pur2 + pur3
authority =~ auth1 + auth2 + auth3
loyalty =~ loy1 + loy2 + loy3 
ancestors =~ anc1 + anc2 + anc3
'
#much better
cfa_modN2 = cfa(itemsN, data = QsN)
summary(cfa_modN2, standardized = TRUE, fit.measures=TRUE)
```

Even better model now those statements have been removed so we can create the indices.

```{r}
get_neg_morals = function(negative){
  negative$compassion = rowMeans(data.frame(negative$Q6_1_a, negative$Q6_2_a,negative$Q6_3_a), na.rm = TRUE)

  negative$fairness = rowMeans(data.frame(negative$Q6_5_a,negative$Q6_6_a), na.rm = TRUE)

  negative$purity = rowMeans(data.frame(negative$Q7_2_a, negative$Q7_3_a), na.rm = TRUE)

  negative$authority = rowMeans(data.frame(negative$Q7_4_a, negative$Q7_5_a, negative$Q7_6_a), na.rm = TRUE)

  negative$loyalty = rowMeans(data.frame(negative$Q8_1_a, negative$Q8_2_a, negative$Q8_3_a), na.rm = TRUE)

  negative$ancestors = rowMeans(data.frame(negative$Q8_4_a, negative$Q8_5_a, negative$Q8_6_a), na.rm = TRUE)

  return(negative)
}

negative = get_neg_morals(negative)

```


Now repeat the process for positive statements
```{r}
QsP= data.frame(positive$Q6_1_a, positive$Q6_2_a,positive$Q6_3_a,positive$Q6_4_a, positive$Q6_5_a,positive$Q6_6_a, positive$Q7_1_a, positive$Q7_2_a, positive$Q7_3_a, positive$Q7_4_a, positive$Q7_5_a, positive$Q7_6_a , positive$Q8_1_a, positive$Q8_2_a, positive$Q8_3_a, positive$Q8_4_a, positive$Q8_5_a, positive$Q8_6_a)

names(QsP) = c("comp1", "comp2", "comp3", "fair1", "fair2", "fair3", "pur1", "pur2", "pur3", "auth1", "auth2", "auth3", "loy1", "loy2", "loy3", "anc1", "anc2", "anc3")

items = '
compassion =~ comp1 + comp2 + comp3
fairness =~ fair1 + fair2 + fair3
purity =~ pur1 + pur2 + pur3
authority =~ auth1 + auth2 + auth3
loyalty =~ loy1 + loy2 + loy3 
ancestors =~ anc1 + anc2 + anc3
'

#pretty bad model but can also be improved
cfa_modP = cfa(items, data = QsP)
summary(cfa_modP, standardized = TRUE, fit.measures=TRUE) 

```

Remove the statements with low loadings and run again

```{r}
#remove comp1, pur2, loy1, fair3
itemsP = '
compassion =~ comp2 + comp3
fairness =~ fair1 + fair2 
purity =~ pur1 + pur3
authority =~ auth1 + auth2 + auth3
loyalty =~ loy2 + loy3 
ancestors =~ anc1 + anc2 + anc3
'
#a bit better but still under the thresholds
cfa_modP2 = cfa(itemsP, data = QsP)
summary(cfa_modP2, standardized = TRUE, fit.measures=TRUE) 
```
The positive model is still under the recommended thresholds so these indices do not work for the positive statements. Instead of indices we'll just use the most convincing statement from each moral category. 

```{r}
names(positive)[names(positive) == "Q6_2_a"] <- "CP2"
names(positive)[names(positive) == "Q6_5_a"] <- "FP2"
names(positive)[names(positive) == "Q7_3_a"] <- "PP3"
names(positive)[names(positive) == "Q7_6_a"] <- "AP3"
names(positive)[names(positive) == "Q8_1_a"] <- "LP1"
names(positive)[names(positive) == "Q8_6_a"] <- "GAP3"
```

## Measuring convincincness ratings

Most convincing among the negative indices

```{r warning=FALSE}
library(reshape2)
library(ggplot2)
neg_melt = melt(negative, measure.vars = c("compassion", "fairness", "purity", "authority", "loyalty", "ancestors"), na.rm=TRUE)
ggplot(neg_melt, aes(x = variable, y = value)) + geom_boxplot() + ggtitle("Negative Statements, Indices") + xlab("Moral Category") + ylab("Convincing") + scale_y_continuous(breaks = c(1,2,3,4,5,6))
```

And the same for mot convincing positive statements
```{r warning=FALSE}
pos_melt = melt(positive, measure.vars = c("CP2", "FP2", "PP3", "AP3", "LP1", "GAP3"), na.rm=TRUE)
ggplot(pos_melt, aes(x = variable, y = value)) + 
  geom_boxplot() +stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red", fill="red") + 
  ggtitle("Positive Statements, Single") + xlab("Moral Category") + ylab("Convincing") + 
  scale_y_continuous(breaks = c(1,2,3,4,5,6))
```

Recode the applicability and novelty questions so they are 0 and 1.

```{r}
applicableN = data.frame(negative$Q6_1_b, negative$Q6_2_b,negative$Q6_3_b,negative$Q6_4_b, negative$Q6_5_b,negative$Q6_6_b, negative$Q7_1_b, negative$Q7_2_b, negative$Q7_3_b, negative$Q7_4_b, negative$Q7_5_b, negative$Q7_6_b , negative$Q8_1_b, negative$Q8_2_b, negative$Q8_3_b, negative$Q8_4_b, negative$Q8_5_b, negative$Q8_6_b)
applicableN[applicableN == 2] <- 0 #not applicable
applicableN[applicableN == 1] <- 1 #applicable


applicableP = data.frame(positive$Q6_1_b, positive$Q6_2_b,positive$Q6_3_b,positive$Q6_4_b, positive$Q6_5_b,positive$Q6_6_b, positive$Q7_1_b, positive$Q7_2_b, positive$Q7_3_b, positive$Q7_4_b, positive$Q7_5_b, positive$Q7_6_b , positive$Q8_1_b, positive$Q8_2_b, positive$Q8_3_b, positive$Q8_4_b, positive$Q8_5_b, positive$Q8_6_b)
applicableP[applicableP == 2] <- 0 #not applicable
applicableP[applicableP == 1] <- 1 #applicable


##Novel (closer to 2 = more novel)
novelsN = negative[, 66:85]
novelsN = novelsN[, -c(7,14)]

novelsN[novelsN == 1] <- 0 #not novel
novelsN[novelsN == 2] <- 1 #novel
novelsN = data.frame(novelsN)


novelsP = positive[ , 66:85]
novelsP = novelsP[ , -c(7,14)]

novelsP[novelsP == 1] <- 0 #not novel
novelsP[novelsP == 2] <- 1 #novel
```


## Correlations

Calculate correlations between convincingness ratings and applicability, and convincingness ratings and novelty

```{r}
pBiserialcorrs = function(x,y){
  corlist = list()
  for(i in 1:18){
    check = cor.test(as.numeric(unlist(x[,i])), y[,i], alternative = "t")
    
    corlist = append(corlist, c(check$statistic, round(check$p.value, 4),round(check$estimate, 2)))
  }
  
  corframe = data.frame(matrix(unlist(corlist), nrow = 18, byrow = TRUE))
  colnames(corframe) = c("t-statistic", "p-value", "correlation")
  corframe$statement = colnames(y)
  return(corframe)
}
```

Convincingness and novelty
```{r}
#
novelsQsN = pBiserialcorrs(novelsN, QsN)
summary(novelsQsN$correlation)

novelsQsP = pBiserialcorrs(novelsP, QsP)
summary(novelsQsP$correlation)

```

Convincingness and applicability
```{r}
applicableQsN = pBiserialcorrs(applicableN, QsN)
summary(applicableQsN$correlation)

applicableQsP = pBiserialcorrs(applicableP, QsP)
summary(applicableQsP$correlation)

```


Novelty and applicability


```{r}
corrs = function(x,y){
  corlist = list()
  for(i in 1:18){
    check = cor(x[i], y[i], use = "complete.obs")
    
    corlist = append(corlist, round(check[1], 2))
  }
  
  corframe = data.frame(matrix(unlist(corlist), nrow = 18, byrow = TRUE))
  colnames(corframe) = c("correlation")
  corframe$statement = c("comp1", "comp2", "comp3", "fair1", "fair2", "fair3", "pur1", "pur2", "pur3", "auth1", "auth2", "auth3", "loy1", "loy2", "loy3", "anc1", "anc2", "anc3")
  return(corframe)
}
```

```{r}
applicableNovelsN = corrs(applicableN, novelsN)
summary(applicableNovelsN$correlation)

applicableNovelsP = corrs(applicableP, novelsP)
summary(applicableNovelsP$correlation)

```


Calculate correlations between values and morals
```{r message=FALSE, fig.width = 10}
library(psych)
library(corrplot)

correlations_neg = cor(negative[, c("SCH_Conservation","conformity_val", "self_direction_val", "SCH_SelfTrancendence", "universalism_val", "achievement_val", "Political_Leaning", "compassion", "fairness", "purity", "authority", "loyalty", "ancestors", "Income", "Age" )], use = "complete.obs")

corrplot(round(correlations_neg, 2), type = "upper",order = "original", 
         tl.col = "black", tl.srt = 45, method = "number", diag = FALSE, number.cex = 0.8, tl.cex = 0.9, cl.cex = 0.8)

```

Repeat for positive statements

```{r, fig.width = 10}
correlations_pos = cor(positive[, c("SCH_Conservation","conformity_val", "self_direction_val", "SCH_SelfTrancendence", "universalism_val", "achievement_val", "Political_Leaning", "CP2", "FP2", "PP3", "AP3", "LP1", "GAP3", "Income", "Age" )], use = "complete.obs")

corrplot(round(correlations_pos, 2), type = "upper",order = "original", 
         tl.col = "black", tl.srt = 45, method = "number", diag = FALSE, number.cex = 0.8, tl.cex = 0.9, cl.cex = 0.8)

```


## T-tests

Calculate t-tests to compare how different groups were convinced by statements


Political leaning
```{r message=FALSE}
library(dplyr)
negative = negative %>% 
  mutate(politics = case_when(Political_Leaning < 5 ~ "Left", Political_Leaning >= 5 ~ "Right") )

table(negative$politics)

leftN = subset(negative, politics == "Left")
rightN = subset(negative, politics == "Right")    

t.test(leftN$compassion, rightN$compassion, var.equal = FALSE)
t.test(leftN$fairness, rightN$fairness, var.equal = FALSE)
t.test(leftN$purity, rightN$purity, var.equal = FALSE)
t.test(leftN$authority, rightN$authority, var.equal = FALSE)
t.test(leftN$loyalty, rightN$loyalty, var.equal = FALSE)
t.test(leftN$ancestors, rightN$ancestors, var.equal = FALSE)
```

Signifiant p-values so can visualise in a boxplot
```{r}
neg_melt = melt(negative, measure.vars = c("compassion", "fairness", "purity", "authority", "loyalty", "ancestors"), na.rm=TRUE)

#Which political leaning of participants are more convinced?
ggplot(neg_melt, aes(x = variable, y = value, fill = politics)) + geom_boxplot() + xlab("Moral Category") + ylab("Convincing") + scale_y_continuous(breaks = c(1,2,3,4,5,6))

```

Repeat for positive statements but they are all insignificant
```{r}
positive = positive %>% 
  mutate(politics = case_when(Political_Leaning < 5 ~ "Left", Political_Leaning >= 5 ~ "Right") )

pos_melt = melt(positive, measure.vars = c("CP2", "FP2", "PP3", "AP3", "LP1", "GAP3"), na.rm=TRUE)

#Which political leaning of participants are more convinced?
ggplot(pos_melt, aes(x = variable, y = value, fill = politics)) + geom_boxplot() + xlab("Moral Category") + ylab("Convincing") + scale_y_continuous(breaks = c(1,2,3,4,5,6))


table(positive$politics)

leftP = subset(positive, politics == "Left")
rightP = subset(positive, politics == "Right")    

#insignificant
t.test(leftP$CP2, rightP$CP2, var.equal = FALSE)
t.test(leftP$FP2, rightP$FP2, var.equal = FALSE)
t.test(leftP$PP3, rightP$PP3, var.equal = FALSE)
t.test(leftP$AP3, rightP$AP3, var.equal = FALSE)
t.test(leftP$LP1, rightP$LP1, var.equal = FALSE)
t.test(leftP$GAP3, rightP$GAP3, var.equal = FALSE)

```

Academic background

```{r}
#negative
negative = negative %>% 
  mutate(academics = case_when(Q15 < 5 ~ "non-academic", Q15 >= 5 ~ "academic") )

table(negative$academics)

aca = subset(negative, academics == "academic")
nonaca = subset(negative, academics == "non-academic")   

#non signif
t.test(aca$compassion, nonaca$compassion, var.equal = FALSE)
t.test(aca$fairness, nonaca$fairness, var.equal = FALSE)
t.test(aca$purity, nonaca$purity, var.equal = FALSE)
t.test(aca$authority, nonaca$authority, var.equal = FALSE)
t.test(aca$loyalty, nonaca$loyalty, var.equal = FALSE)
t.test(aca$ancestors, nonaca$ancestors, var.equal = FALSE)

```


```{r}
#positive
positive = positive %>% 
  mutate(academics = case_when(Q15 < 5 ~ "non-academic", Q15 >= 5 ~ "academic") )

table(positive$academics)

acaP = subset(positive, academics == "academic")
nonacaP = subset(positive, academics == "non-academic")   
#insignificant
t.test(acaP$CP2, nonacaP$CP2, var.equal = FALSE)
t.test(acaP$FP2, nonacaP$FP2, var.equal = FALSE)
t.test(acaP$PP3, nonacaP$PP3, var.equal = FALSE)
t.test(acaP$AP3, nonacaP$AP3, var.equal = FALSE)
t.test(acaP$LP1, nonacaP$LP1, var.equal = FALSE)
t.test(acaP$GAP3, nonacaP$GAP3, var.equal = FALSE)
```


Religion

Just perform t-test for the religiously-framed statements
```{r}
negative = negative %>% 
  mutate(religion = case_when(Q18 < 9 ~ "religious", Q18 == 9 ~ "non-religious") )

table(negative$religion)

relig = subset(negative, religion == "religious")
nonrelig = subset(negative, religion == "non-religious")   
#religious statement
t.test(relig$Q7_1_a, nonrelig$Q7_1_a, var.equal = FALSE)
```

And repeat for positive

```{r}
#positive
positive = positive %>% 
  mutate(religion = case_when(Q18 < 9 ~ "religious", Q18 == 9 ~ "non-religious") )

table(positive$religion)

religP = subset(positive, religion == "religious")
nonreligP = subset(positive, religion == "non-religious") 

#religious statement
t.test(religP$Q7_2_a, nonreligP$Q7_2_a, var.equal = FALSE)
```

